---
title: "ggplot2 Hands-On Practice"
format: html
---

```{r setup, include=FALSE}
library(tidyverse)
library(palmerpenguins)
```

Time to apply what we learned! Work through the activity that matches your experience level. If you finish early, try the bonus question or move on to the next level!

---

## Activity 1: Fill in the blanks

**Goal:** Build a scatterplot of penguin bill length vs. bill depth, colored by
species, using what you learned today.

**Instructions:** Fill in the `___` blanks to complete the code. Run it and
check that your plot displays correctly!

```{r beginner, eval=FALSE}
ggplot(
  ___ = penguins,
  mapping = aes(x = bill_length_mm, y = ___, color = ___)
) +
  ___() +
  labs(
    title = "Bill length and bill depth by species",
    x = "Bill length (mm)",
    y = "___",
    color = "___"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(___ = 0.5), 
    plot.subtitle = element_text(___ = 0.5)
  )
```

**Hints:**

- The dataset you've been using all day goes in the first blank
- Find the variable name for bill depth
- Use the same geom as the scatterplots in the presentation
- Axis labels should be human-readable (e.g., "Bill depth (mm)")
- Revist **Manually modifying theme** slides to center titles

**Bonus:** Can you also add `shape = species` inside `aes()` to make the
plot accessible to colorblind viewers?

---

## Activity 2: Update the facet plots

**Goal:** Recreate and then enhance the faceted penguin plot from the
presentation.

**Instructions:** Start with the faceted plot from the slides, then make the
following modifications:

1. **Apply a theme** — add a theme with the following specifications:
- Position the title and subtitle to be right aligned 
- Set the legend position to "bottom"
- Remove major and minor gridlines
- Add axes lines that are colored blue
- Set your theme as a function before applying it to your figure.
2. **Add color** — map `color = species` inside so each species
panel has a distinct color
3. **Fix the legend position** — the default `legend.position = c(0.85, 0.2)`
may overlap the facets; adjust it to `"bottom"` instead
**Bonus:** Try `facet_grid(sex ~ species)` instead of `facet_wrap(~ species)`.
What changes? What does this tell you that the original plot didn't?

```{r intermediate, eval=FALSE}
# Facet plots from presentation - modify according to instructions above
  ggplot(
    data = penguins,
    mapping = aes(x = flipper_length_mm, y = body_mass_g)
  ) +
    geom_point() +
    facet_wrap(~ species) +
    labs(
      title = "Penguin body mass by flipper length",
      subtitle = "Faceted by species",
      x = "Flipper length (mm)",
      y = "Body Mass (g)"
    )
```

---


## Activity 3: Create figure using new dataset

**Goal:** Use the `mtcars` dataset to independently apply everything from
the presentation to answer a research question.

**Background:** `mtcars` is a built-in R dataset containing performance and
design specs for 32 car models. Run `?mtcars` to see the full variable list.
Key variables:

| Variable | Description |
|----------|-------------|
| `mpg` | Miles per gallon |
| `hp` | Horsepower |
| `wt` | Weight (1000 lbs) |
| `cyl` | Number of cylinders (4, 6, or 8) |
| `am` | Transmission (0 = automatic, 1 = manual) |

**Research question:** *Does the relationship between horsepower and fuel
efficiency differ by number of cylinders?*

**Your task — build a complete, publication-ready figure that:**

1. Plots horsepower (x) vs. mpg (y) as a scatterplot
2. Assigns colors and shapes points by number of cylinders
 — the variable is numeric, so you'll need `factor(cyl)` inside `aes()` to treat it as a category
3. Adds a line of best fit **per cylinder group** `
4. Facets by transmission type 
— use `labeller = labeller(am = c("0" = "Automatic", "1" = "Manual"))` inside
`facet_wrap()` to replace 0/1 with readable labels
5. Applies a clean custom theme
6. Has polished axis labels, a title, and a subtitle

```{r advanced, eval=FALSE}
# Your code here!
```

**Bonus:** Add `se = TRUE` back to `geom_smooth()` (it's the default). What
does the shaded ribbon represent, and when might it be useful in a
scientific figure?
